---
title: "Mini Data Analysis Milestone 2"
output: github_document
---

*To complete this milestone, you can either edit [this `.rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the output in the YAML header to `github_document`, before submitting a tagged release on canvas.*

# Welcome back to your mini data analysis project!

This time, we will explore more in depth the concept of *tidy data*, and hopefully investigate further into your research questions that you defined in milestone 1.

**NOTE**: The main purpose of the mini data analysis is to integrate what you learn in class in an analysis. Although each milestone provides a framework for you to conduct your analysis, it's possible that you might find the instructions too rigid for your data set. If this is the case, you may deviate from the instructions -- just make sure you're demonstrating a wide range of tools and techniques taught in this class.

Begin by loading your data and the tidyverse package below: 

```{r, message = FALSE}
library(datateachr) # <- might contain the data you picked!
suppressMessages(library(tidyverse))
suppressMessages(library(ggridges))
```

# Learning Objectives

By the end of this milestone, you should:

+ Become familiar with manipulating and summarizing your data in tibbles using `dplyr` and `tidyr`, with a research question in mind.
+ Understand what *tidy* data is, and how to create it. In milestone 3, we will explore when this might be useful.
+ Generate a reproducible and clear report using R Markdown.
+ Gain a greater understanding of how to use R to answer research questions about your data.

**Things to keep in mind**

+ Remember to document your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. Create your analysis as if someone else will be reading it! **There will be 2.5 points reserved for reproducibility, readability, and repo organization.**

+ Before working on each task, you should always keep in mind the specific **research question** that you're trying to answer.

# Task 1: Process and summarize your data (15 points)

From milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions. 

### 1.1 (2.5 points) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->
1. *In what years was the central air-conditioning system mostly installed in buildings?*
2. *Which heating type is the most common among very tall buildings?*
3. *Is there a change in the number of storeys built over the years in relation to window type?*
4. *What is the relationship between number of accessible parking spaces and number of storeys for each visitor parking category by property type?*

*I added categorical variables to question 4 since there's no exercise given for investigating the correlation between numerical variables only*
<!----------------------------------------------------------------------------->

### 1.2 (10 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question. 

Ensure that the output of each operation is printed! 

**Summarizing:**

1. Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2. Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3. Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Based on two categorical variables, calculate two summary statistics of your choosing.

**Graphing:**

5. Create a graph out of summarized variables that has at least two geom layers.
6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read. 
7. Make a graph where it makes sense to customize the alpha transparency.
8. Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->
**Research Question 1: In what years was the central air-conditioning system mostly installed in buildings?**

To answer the question, it would be good to determine what the air-conditioning types are and their frequencies especially that of the *central air*. Then, we would go on to visualize the frequency of use throughout the years under review(1805 - 2019).

```{r}
# Compute the number of observations for at least one of your categorical variables(air_conditioning)
apt_buildings %>% 
  group_by(air_conditioning) %>% 
  summarize(n = n())

# Make a graph where it makes sense to customize the alpha transparency: determine when central air-conditioning system was mostly installed

apt_buildings %>% 
  filter(!is.na(year_built), !is.na(air_conditioning)) %>% 
  ggplot(aes(year_built, fill = air_conditioning)) + 
  geom_density(alpha = 0.4)

```


In trying to answer **research question 1 above**, it made sense to first determine the total number of `central air` air-conditioning units installed in buildings. Surprisingly, it was discovered that `most of the buildings have no air-conditioning installed!` `Eighty-five` buildings have missing information on air-conditioning.

To further answer the research question, a density plot was used to determine when central air-conditioning was mostly installed in buildings. The plot clearly lays out the 3 air-conditioning systems where it shows that `central air` installation increased after 1950 and remained the most installed air-conditioning system from around 1975 to 2019.



**Research question 2: Which heating type is the most common among very tall buildings?**

Let me first check the distribution of the different heating types among all the buildings. After that, I will go ahead and visualize that distribution graphically. Hopefully we would be able to see the heating type that is most common among very tall buildings.

```{r, message = FALSE}
# Compute the range, mean, median and sum of one numerical variable(no_of_storeys) across the groups of one categorical variable(heating_type) from your data

apt_buildings %>%
  filter(!is.na(heating_type), !is.na(no_of_storeys)) %>%
  group_by(heating_type) %>%
  summarise(storey_range = paste(min(no_of_storeys), "-", max(no_of_storeys)),
            storey_mean = mean(no_of_storeys, na.rm = TRUE),
            storey_median = median(no_of_storeys, na.rm = TRUE),
            n = sum(!is.na(no_of_storeys)))


# Create a graph out of summarized variables that has at least two geom layers
# Create a boxplot to show range(minimum, maximum), median no_of_storeys across the heating types. Add mean as a second layer 
apt_buildings %>% 
  filter(!is.na(heating_type), !is.na(no_of_storeys)) %>% 
  ggplot(aes(heating_type, no_of_storeys)) + 
  geom_boxplot(aes(fill = heating_type), show.legend = FALSE) + 
  stat_summary(fun = mean, shape = 3, na.rm = TRUE)

```


For **research question 2 above**, first it is shown that `HOT WATER` is the most used heating type, and that 50% of the buildings have 5 - 6 storeys. The average number of storeys to use `ELECTRIC` is approximately 10, the average number of storeys to use `FORCED AIR GAS` is approximately 8, while the average number of storeys to use `HOT WATER` is approximately 8. The tallest buildings to use `ELECTRIC` heating type have `33 storeys`, the tallest buildings to use `FORCED AIR GAS` have `43 storeys`, while the tallest building to use `HOT WATER` has `51 storeys`. 

Graphically, it is seen that `HOT WATER` heating type was installed in the tallest building. However, it is also seen that `FORCED AIR GAS`is fairly common in very tall buildings as well.



**Research question 3: Is there a change in the number of storeys built over the years in relation to window type?**

To answer this question, let me first of all group the years in which the buildings were erected into 5 categories: `ancient(before 1850)`, `very old(before 1900)`, `old(before 1950)`, `recent(before 2000)`, `more recent(before 2020)`. After that, geom_density_ridges will be used to visualize the changes in number of storeys built within the year-groups in relation to window type.

```{r, message = FALSE}
# Create a categorical variable(age_of_building) with 5 groups from a numerical variable(year_built)
(apt_buildings_new <- apt_buildings %>% 
   select(year_built, no_of_storeys, window_type) %>% 
   filter(!is.na(year_built), !is.na(no_of_storeys), !is.na(window_type)) %>% 
   mutate(age_of_building = case_when(year_built < 1850 ~ "ancient", 
                                      year_built < 1900 ~ "very old", 
                                      year_built < 1950 ~ "old", 
                                      year_built < 2000 ~ "recent", 
                                      TRUE ~ "more recent")))

# Make a graph where it makes sense to customize the alpha transparency: visualize the changes in number of storeys built over the years in relation to window type

suppressMessages(print(apt_buildings_new %>% 
  ggplot(aes(no_of_storeys, age_of_building)) + 
  geom_density_ridges(aes(fill = window_type), alpha = 0.6, show.legend = FALSE) + 
  facet_wrap(~ window_type)))

```


I set the alpha transparency of the plot above because some peaks ran into the ones above them. From the plot it is seen that houses with more storeys were built within the `recent` and ` more recent` *age_of_building* groups. `ancient` and ` very old` buildings(erected between 1805 and 1850) have `DOUBLE PANE` window type while `recent` and `more recent` buildings(erected between 1950 and 2020) with more storeys have `THERMAL` and `SINGLE PANE` window types in addition to `DOUBLE PANE`.



**Research question 4: What is the relationship between number of accessible parking spaces and number of storeys for each visitor parking category by property type?**


```{r, message = FALSE}
# Compute the number of observations for at least one of your categorical variables (property_type)
apt_buildings %>% 
  group_by(property_type) %>% 
  summarize(n = n())


# Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read
# Visualize the relationship between number of accessible parking spaces and number of storeys for each visitor parking category by property type
apt_buildings %>% 
  filter(!is.na(visitor_parking), 
         !is.na(no_of_storeys), 
         !is.na(no_of_accessible_parking_spaces)) %>% 
  ggplot(aes(no_of_accessible_parking_spaces, no_of_storeys), na.rm = TRUE) + 
  geom_jitter(aes(color = property_type), na.rm = TRUE, width = 0.4, alpha = 0.3) + 
  facet_wrap(~ visitor_parking) + 
  scale_x_log10() + 
  labs(x = "Number of accessible parking spaces", y = "Number of storeys") + 
  theme_minimal()

```


Most of the `FREE` visitor parking is found in `PRIVATE` buildings. Many `PRIVATE` buildings with the lowest number of storeys lack parking spaces for visitors whereas buildings in the`TCHC` group have `BOTH` free and paid `visitor_parking` and also many in the same category have `PAID` visitor_parking. The least `FREE` visitor parking is found among TCHC buildings. `SOCIAL HOUSING` buildings have the least `BOTH` visitor parking. Generally, `TCHC` buildings have the least number of accessible parking spaces while `PRIVATE` buildings have the highest number of accessible parking spaces.
<!----------------------------------------------------------------------------->

### 1.3 (2.5 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?


<!------------------------- Write your answer here ---------------------------->
**Based on the operations completed, all of the research questions have been answered.** Research question 1 yielded interesting results; it was discovered that majority of the buildings lack air-conditioning units. It would be interesting to find out the property types that were built without air-conditioning system.

```{r}
# Determine property types built without air-conditioning units installed
apt_buildings %>%  filter(!is.na(property_type), 
                          !is.na(no_of_storeys), 
                          !is.na(air_conditioning)) %>% 
  ggplot(aes(air_conditioning, no_of_storeys)) + 
  geom_jitter(width = 0.3, alpha = 0.2) + 
  facet_wrap(~ property_type) +
  coord_flip()
```

The analysis above shows that `PRIVATE` property type has the highest concentration of buildings without air-conditioning system. It also reveals that most `TCHC` buildings don't have air-conditioning system while the few that have make use of the `CENTRAL AIR`. `alpha transparency` and `jittering` were used to reveal overlapping points.

<!----------------------------------------------------------------------------->

# Task 2: Tidy your data (12.5 points)

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

- Each row is an **observation**
- Each column is a **variable**
- Each cell is a **value**

*Tidy'ing* data is sometimes necessary because it can simplify computation. Other times it can be nice to organize data so that it can be easier to understand when read manually. 

### 2.1 (2.5 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have >8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->

**I will extract 8 variables for this question**

```{r}
# Subset the `apt_buildings` dataset: select 8 variables
(apt_buildings_subvar <- apt_buildings %>% 
   select(id, air_conditioning, 
          heating_type, 
          no_of_storeys, 
          visitor_parking, 
          window_type,
          property_type, 
          year_built)
 )
```

The data subset above contains 8 variables and 3455 observations. The `id` variable uniquely identifies each observation. Each row is a single observation - contains all values measured on a particular building, each column is a single variable - contains values that measure the same underlying attribute, each cell is a single value. 

Additionally, considering my research questions, the data is tidy.
<!----------------------------------------------------------------------------->

### 2.2 (5 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".


<!--------------------------- Start your work below --------------------------->
**First, let me show the `before` state of the `apt_buildings_subvar` data in its `tidy` form as already explained above**

```{r}
# The data in a tidy form
(apt_buildings_subvar <- apt_buildings %>% 
   select(id, air_conditioning, 
          heating_type, 
          no_of_storeys, 
          visitor_parking, 
          window_type, 
          property_type, 
          year_built)
 )
```



**Second, below is the state of the `apt_buildings_subvar` data in its *untidy* form**

```{r}
# The data in an untidy form
(apt_buildings_subvar_wider <- apt_buildings_subvar %>% 
   mutate(property_type_val = TRUE) %>%
   pivot_wider(id_cols = c(-property_type, - property_type_val), 
               names_from = property_type, 
               values_from = property_type_val, 
               values_fill = list(property_type_val = FALSE)) %>% 
   select(id, PRIVATE, TCHC, `SOCIAL HOUSING`, everything()))
  
```

As can be seen above, the three new columns `PRIVATE`, `TCHC`, and `SOCIAL HOUSING` are all values that can be grouped under a single variable. For example,  `PRIVATE` makes no meaning as a variable. Besides, in this format, it is untidy to use to answer my research questions. The 3 new columns can be conveniently grouped under a variable that measures the type of property each building belongs to.


**Now, let me revert the `apt_buildings_subvar` data to its original tidy state!**

```{r}
(apt_buildings_subvar_longer <- apt_buildings_subvar_wider %>% 
    pivot_longer(PRIVATE:`SOCIAL HOUSING`, names_to = "property_type", values_to = "property_type_val") %>% 
    filter(property_type_val) %>% 
    select(-property_type_val))
```


The `apt_buildings_subvar` dataset named `apt_buildings_subvar_longer` now has the `property_type` variable measuring the same attribute using the `PRIVATE`, `TCHC`, and `SOCIAL HOUSING` values. It is now in the original tidy state.

<!----------------------------------------------------------------------------->

### 2.3 (5 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in milestone 3, and explain your decision. 

Try to choose a version of your data that you think will be appropriate to answer these 2 questions in milestone 3. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelvant columns, etc.). 

<!--------------------------- Start your work below --------------------------->
**Two research questions for analysis in milestone 3**

1. Which heating type is the most common among very tall buildings?
2. Is there a change in the number of storeys built over the years in relation to window type?

I chose the 2 questions above because of their relevance to tasks in milestone 3 which involve reordering, grouping factors, model-fitting, etc. They produce plots that have at least 3 groups. All these activities require tidy data and I know that the data that generated the plots is tidy.


```{r}
# Choose a version of the apt_buildings data appropriate to answer the 2 research questions for milestone 3
(apt_buildings %>% 
   select(id, heating_type, no_of_storeys, year_built, window_type) %>%
   filter(!is.na(heating_type), !is.na(no_of_storeys), !is.na(year_built),!is.na(window_type)) %>%
   mutate(age_of_building = case_when(year_built < 1850 ~ "ancient", 
                                      year_built < 1900 ~ "very old", 
                                      year_built < 1950 ~ "old", 
                                      year_built < 2000 ~ "recent", 
                                      TRUE ~ "more recent")) %>%
   mutate(across(where(is.character), tolower)) %>%
   arrange(desc(year_built))
 )
```

First, I piped the apt_buildings data into the `select` function where I selected five variables that will be appropriate for milestone 3. Next, I piped it into the `filter` function to remove all missing values in the selected variables. I piped that again into the `mutate` function where I created a new variable named `age_of_building` which grouped the years that the buildings were erected into 5 categories. The variable names are in lowercase but the character values are in uppercase. So, I cleaned the dataset by converting every character to lowercase for uniformity. Finally, I rearranged the dataset by arranging the `year_built` variable in descending order.
<!----------------------------------------------------------------------------->

*When you are done, knit an `md` file. This is what we will mark! Make sure to open it and check that everything has knitted correctly before submitting your tagged release.*

### Attribution

Thanks to Victor Yuan for mostly putting this together. 