---
title: "Mini Data-Analysis Deliverable 1"
date: "October 9, 2021"
author: "Ekpereka Amutaigwe"
output: 
   github_document:
      toc: true
      toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## 1. Introduction

This mini-project is part of the deliverables in the STAT545A course. It involves installation of the `datateachr` package by Hayley Boyce and Jordan Bourak which currently consists of 7 semi-tidy datasets.

It also requires loading the `datateachr` and `tidyverse` packages which are necessary for a successful completion of the project.


**There are 3 major tasks in this project**:

+ Choose your favorite dataset
+ Explore your dataset
+ Write your research questions


### 1.1 Learning Objectives

By the end of this project, I should:

+ Become familiar with the datasets, especially the one I choose to work with in greater depth
+ Be able to formulate 4 research questions that I would like to answer with my data
+ Generate a well written and reproducible report using R Markdown


### 1.2 Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package 

To install the **datateachr** package, first install the **devtools** package using **install.packages()** with parentheses surrounding *devtools* in your **R console** and load the **devtools** package using **library()** before installing *datateachr* with this code-  **devtools::install_github("UBC-MDS/datateachr")**. It is a one-off exercise in that once it is installed, you only need to reload it whenever you start a new session to use the package and the datasets it contains.


### 1.3 Load required packages

```{r}
library(datateachr)
library(tidyverse)
```



## 2. Define variables

It is always beneficial to define the variables/datasets you will be using in your analysis early in your project. The datasets can be explored without defining them once the package containing them is loaded. However, for clarity and to notice at a glance the source of the datasets, I like to define them early enough.

```{r}
# define variables
apt_buildings <- datateachr::apt_buildings
building_permits <- datateachr::building_permits
cancer_sample <- datateachr::cancer_sample
flow_sample <- datateachr::flow_sample
parking_meters <- datateachr::parking_meters
steam_games <- datateachr::steam_games
vancouver_trees <- datateachr::vancouver_trees
```



## 3. Task 1: Choose your favorite dataset

As mentioned earlier, the following are the 7 datasets to choose from:

+ *apt_buildings*: Acquired courtesy of The City of Toronto's Open Data Portal. It currently has 3455 rows and 37 columns.
+ *building_permits*: Acquired courtesy of the City of Vancouver's Open Data Portal. It currently has 20680 rows and 14 columns.
+ *cancer_sample*: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.
+ *flow_sample*: Acquired courtesy of The Government of Canada's Historical Hydrometric Database. It currently has 218 rows and 7 columns.
+ *parking-meters*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 10032 rows and 22 columns.
+ *steam_games*: Acquired courtesy of Kaggle. It currently has 40833 rows and 21 columns.
+ *vancouver_trees*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 146611 rows and 20 columns.


**The above description is an excerpt from the milestone document for the project prepared by Icíar Fernández Boyano and launched by Vincenzo Coia**.



### 3.1. Instructions for the project as given by the teaching team

**Instruction 1.1**: Out of the 7 datasets available in the **datateachr** package, choose **4** that appeal to you based on their description.


#### 3.1.1. The 4 choices

1. apt_buildings
2. cancer_sample
3. flow_sample
4. vancouver_trees


**Instruction 1.2**: One way to narrow down your selection is to *explore* the datasets. Use your knowledge of dplyr to find out at least 3 attributes about each of these datasets. The goal is to have an idea of *what the data looks like*


#### 3.1.2. Exploration of the datasets to have an idea of what they look like

To explore each of the four chosen datasets, I will look at class of the data, variable types, number of rows and columns, the variable names, the total number of missing values (NAs). 

**The code below are from what I learned in STAT545A and the R for Data Science book by Hadley Wickham & Garrett Grolemund.**



```{r}
# explore the apt_buildings dataset
# get class of the dataset
class(apt_buildings)

# find out number of rows and columns, names of variables, variable types
glimpse(apt_buildings)

# get total number of missing values(NAs)
sum(is.na(apt_buildings))

```


From the analysis above, one can conclude that the *apt_buildings* is a dataframe that has been converted to a tibble, and has 3,455 rows and 37 columns. The data contains mostly character variables with few doubles. A total of 6286 values are missing (NAs). 




```{r}
# explore the cancer_sample dataset
# get class of the dataset
class(cancer_sample)

# find out number of rows and columns, names of variables, variable types
glimpse(cancer_sample)

# get total number of missing values(NAs)
sum(is.na(cancer_sample))
```


The analysis above shows that the *cancer_sample* dataset is a dataframe that has been converted to a tibble, and has 569 rows and 32 columns. Most of the variables are doubles with only 1 character variable. Surprisingly, there's no missing value in the dataset.




```{r}
# explore the flow_sample dataset
# get class of the dataset
class(flow_sample)

# determine number of rows and columns, names of variables, variable types
glimpse(flow_sample)

# get total number of missing values(NAs)
sum(is.na(flow_sample))
```


The *flow_sample* dataset is a dataframe that has been converted to a tibble, and has 218 rows and 7 columns. The variable types are roughly equal- 3 characters and 4 doubles. There are 125 missing values in the dataset.




```{r}
# explore the vancouver_trees dataset
#  get class of the dataset
class(vancouver_trees)

# determine number of rows and columns, names of variables, variable types
glimpse(vancouver_trees)

# get total number of missing values
sum(is.na(vancouver_trees))
```


To summarize findings from the *vancouver_trees* dataset, it is a dataframe that has been converted to a tibble, and comprises 146,611 rows and 20 columns. There are more character variables than doubles. There's one date variable type. There are multiple missing values in the dataset(191,135).



