---
title: "Mini Data-Analysis Deliverable 1"
date: "October 8, 2021"
author: "Ekpereka Amutaigwe"
output: 
   github_document:
      toc: true
      toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## 1. Introduction

This mini-project is part of the deliverables in the STAT545A course. It involves installation of the `datateachr` package by Hayley Boyce and Jordan Bourak which currently consists of 7 semi-tidy datasets.

It also requires loading the `datateachr` and `tidyverse` packages which are necessary for a successful completion of the project.


**There are 3 major tasks in this project**:

+ Choose your favorite dataset
+ Explore your dataset
+ Write your research questions


### 1.1 Learning Objectives

By the end of this project, I should:

+ Become familiar with the datasets, especially the one I choose to work with in greater depth
+ Be able to formulate 4 research questions that I would like to answer with my data
+ Generate a well written and reproducible report using R Markdown


### 1.2 Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package 

To install the **datateachr** package, first install the **devtools** package using **install.packages()** with parentheses surrounding *devtools* in your **R console** and load the **devtools** package using **library()** before installing *datateachr* with this code-  **devtools::install_github("UBC-MDS/datateachr")**. It is a one-off exercise in that once it is installed, you only need to reload it whenever you start a new session to use the package and the datasets it contains.


### 1.3 Load required packages

```{r}
library(datateachr)
library(tidyverse)
```



## 2. Define variables

It is always beneficial to define the variables/datasets you will be using in your analysis early in your project. The datasets can be explored without defining them once the package containing them is loaded. However, for clarity and to notice at a glance the source of the datasets, I like to define them early enough.

```{r}
# define variables
apt_buildings <- datateachr::apt_buildings
building_permits <- datateachr::building_permits
cancer_sample <- datateachr::cancer_sample
flow_sample <- datateachr::flow_sample
parking_meters <- datateachr::parking_meters
steam_games <- datateachr::steam_games
vancouver_trees <- datateachr::vancouver_trees
```



