---
title: "Mini Data-Analysis Deliverable 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## 1. Introduction

This mini-project is part of the deliverables in the STAT545A course. It involves installation of the `datateachr` package by Hayley Boyce and Jordan Bourak which currently consists of 7 semi-tidy datasets.

It also requires loading the `datateachr` and `tidyverse` packages which are necessary for a successful completion of the project.


**There are 3 major tasks in the project**:

+ Choose your favorite dataset
+ Explore your dataset
+ Write your research questions


## 1.1 Learning Objectives

By the end of this project, I should:

+ Become familiar with the datasets, especially the one I choose to work with in greater depth
+ Be able to formulate 4 research questions that I would like to answer with my data
+ Generate a well written and reproducible report using R Markdown


## 1.2 Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package 

Type the following code in your **R terminal**. It is a one-off exercise in that once it is installed, you only need to load it whenever you want to use the package and the datasets it contains.

```
# install package
install.packages("devtools") # load `devtools` before running the next code
devtools::install_github("UBC-MDS/datateachr")
```

## 1.3 Load required packages

```{r}
library(datateachr)
library(tidyverse)
```


## 2. Define variables

It is always beneficial to define the variables/datasets you will be using in your analysis early in your project. The datasets can be explored without defining them once the package containing them is loaded. However, for clarity and to notice at a glance the source of the datasets, I like to define them early enough.

```{r}
# define variables
apt_buildings <- datateachr::apt_buildings
building_permits <- datateachr::building_permits
cancer_sample <- datateachr::cancer_sample
flow_sample <- datateachr::flow_sample
parking_meters <- datateachr::parking_meters
steam_games <- datateachr::steam_games
vancouver_trees <- datateachr::vancouver_trees
```


## 3. Task 1: Choose your favorite dataset

As mentioned earlier, the following are the 7 datasets to choose from:

+ *apt_buildings*: Acquired courtesy of The City of Toronto's Open Data Portal. It currently has 3455 rows and 37 columns.
+ *building_permits*: Acquired courtesy of the City of Vancouver's Open Data Portal. It currently has 20680 rows and 14 columns.
+ *cancer_sample*: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.
+ *flow_sample*: Acquired courtesy of The Government of Canada's Historical Hydrometric Database. It currently has 218 rows and 7 columns.
+ *parking-meters*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 10032 rows and 22 columns.
+ *steam_games*: Acquired courtesy of Kaggle. It currently has 40833 rows and 21 columns.
+ *vancouver_trees*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 146611 rows and 20 columns.


**The above description is an excerpt from the milestone document for the project prepared by Icíar Fernández Boyano and launched by Vincenzo Coia**.


### 3.1. Instructions for the project
1. Out of the 7 datasets available in the **datateachr** package, choose **4** that appeal to you based on their description.

The **4** choices are:

1: apt_buildings
2: cancer_sample
3. flow_sample
4. vancouver_trees

