---
title: "Mini Data_Analysis Deliverable 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

This mini-project is part of the deliverables in the STAT545A course. It involves installation of the `datateachr` package by Hayley Boyce and Jordan Bourak which currently consists of 7 semi-tidy datasets.

It also requires loading the `datateachr` and `tidyverse` packages which are necessary for a successful completion of the project.

There are 3 major tasks in the project:
1. Choose your favorite dataset
2. Explore your dataset
3. Write your research questions

## 1.1 Learning Objectives

By the end of this project, I should:
+ Become familiar with the datasets, especially the one I choose to work with in greater depth
+ Be able to formulate 4 research questions that I would like to answer with my data
+ Generate a well written and reproducible report using R Markdown

## 1.2 Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package 

Type the following code in your **R terminal**. It is a one-off exercise in that once it is installed, you only need to load it whenever you want to use the package and the datasets it contains.

```
# install package
install.packages("devtools") # load `devtools` before running the next code
devtools::install_github("UBC-MDS/datateachr")
```

## 1.3 Load the following packages

```{r}
library(datateachr)
library(tidyverse)
```


